---
title: "Badger Spatial MoL Ecol Resubmission"
author: "Xav Harrison"
date: "2025-06-18"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Libraries and Data

## Libraries
```{r}
  library(phyloseq)
  library(dplyr)
  library(decontam)
  library(ggplot2)
  library(vegan)
  library(ggordiplots)
  library(microbiome)
  library(pheatmap)
  library(cowplot)
  library(RColorBrewer)
  #library(ggvegan)
  library(ggrepel)
  #remotes::install_github("yiluheihei/microbiomeMarker")
  library(microbiomeMarker)
  library(gllvm)
  #devtools::install_github("microsud/microbiomeutilities")
  library(microbiomeutilities)
  #devtools::install_github("BlakeRMills/MoMAColors")
  library(MoMAColors)
  #BiocManager::install("MicrobiotaProcess")
  #library(MicrobiotaProcess)
  library(Polychrome)
  #library(ecodist)
  

```

## Global Plotting Options 
Quick access options for output graphics to make legends and axis labels larger / more legible

```{r}

#Global Plot Options
            plotopts<- theme(axis.text.y=element_text(size=20),axis.text.x=element_text(size=12),axis.title=element_text(size=20),strip.text=element_text(size=20),legend.title = element_text(size=20),legend.text = element_text(size=20)) 

#Smalller Legend Labs
            plotopts2<- theme(axis.text.y=element_text(size=20),axis.text.x=element_text(size=12),axis.title=element_text(size=20),strip.text=element_text(size=20),legend.title = element_text(size=15),legend.text = element_text(size=12)) 

## Global Site Colors
  sitecols<-c(brewer.pal(8,"Set2"),brewer.pal(9,"Paired")[9])


###### Stop Warnings turning into Errors
  options(warn=1)
```

## Microbiome Data
```{r}

  ps_woodchester<-readRDS('Woodchester Phyloseq Jun25 RID.Rdata')

```

# Decontamination 

## Data Cleanup 
```{r}

############ DATA CLEANUP 
  
  #Prune Taxa With No Phylum Assignment 
    ps_prune<-prune_taxa(as.vector(!is.na(tax_table(ps_woodchester)[,2])),ps_woodchester)
      ntaxa(ps_woodchester)-ntaxa(ps_prune) 
      
  #Prune Chloroplasts
      ps_prune_nochloro<-prune_taxa(as.vector(tax_table(ps_prune)[,4]!="Chloroplast"),ps_prune)
      ntaxa(ps_prune)-ntaxa(ps_prune_nochloro) 
      
  #Remove Mitochondria    
      ps_prune_nochloro_nomito<-prune_taxa(as.vector(tax_table(ps_prune_nochloro)[,5]!="Mitochondria"),ps_prune_nochloro)
      
  #Remove Archaea  
    ps_prune_nochloro_nomito_noarchaea<-prune_taxa(as.vector(tax_table(ps_prune_nochloro_nomito)[,1]!="Archea"),ps_prune_nochloro_nomito)
    
     ntaxa(ps_woodchester) - ntaxa(ps_prune_nochloro_nomito_noarchaea)


```

## Plot of Library Sizes
```{r}

      #Inspect Library Sizes
            df <- as.data.frame(sample_data(ps_prune_nochloro_nomito_noarchaea)) # Put sample_data into a ggplot-friendly data.frame
            df$LibrarySize <- sample_sums(ps_prune_nochloro_nomito_noarchaea)
            df <- df[order(df$LibrarySize),]
            df$Index <- seq(nrow(df))
            df$sample_or_control<-ifelse(df$sampletype %in% c("negative","negative_ve"),"negative","sample")
           
            ggplot(data=df, aes(x=Index, y=LibrarySize,colour=sample_or_control)) + geom_point()
        
        
          
```  

## Identify Contaminants

```{r}
        ##Identify Contaminants by Prevalence
            sample_data(ps_prune_nochloro_nomito_noarchaea)$is.neg <- sample_data(ps_prune_nochloro_nomito_noarchaea)$sampletype == "negative"
            contamdf.prev <- isContaminant(ps_prune_nochloro_nomito_noarchaea, method="prevalence", neg="is.neg",threshold=0.6,normalize=T)
            
          #How Many Contaminants?  
            table(contamdf.prev$contaminant) #722
            
```

## Plot of Contaminant Frequency by Sample Type
Seems to be one ASV present in 3 negatives at high abundance - but decontam thinks this is contamination from true samples. Probably because it is not in the other 3 negatives
```{r}

# Make phyloseq object of presence-absence in negative controls and true samples
  ps.pa <- transform_sample_counts(ps_prune_nochloro_nomito_noarchaea, function(abund) 1*(abund>0))
  ps.pa.neg <- prune_samples(sample_data(ps.pa)$is.neg == TRUE, ps.pa)
  ps.pa.pos <- prune_samples(sample_data(ps.pa)$is.neg == FALSE, ps.pa)

# Make data.frame of prevalence in positive and negative samples
  df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)

#Plot
  ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")

```

### Remove Contaminants
```{r}
    ## Remove Contaminants
            ps_clean<-prune_taxa(!contamdf.prev$contaminant,ps_prune_nochloro_nomito_noarchaea)

     ##Remove Mocks and Negatives
            table(sample_data(ps_clean)$sampletype)
            ps_badger<-subset_samples(ps_clean,sampletype=="badger" & !Sample %in% c("A","B"))
```

## Filter Based on Prevalence and Abundance
```{r}

####### Read Threshold and Sample Prevalence Threshold
  ps_clean_filter = filter_taxa(ps_badger, function(x) sum(x) > 10, TRUE)
  #ps_clean_filter<-ps_badger
```


## Assessing Post-QC Sample Coverage and Library Sizes
  
  Now that we've done that, we can check what our post-QC library sizes are. It's a good idea to report these in manuscripts and thesis chapters. Here was ask what the minimum, maximum and mean library sizes-per-samples are. 
  

```{r}
 ############## POST QC LIBRARY SIZES
      
      ###############
      # Post-QC Library Stats
      ###############
      
      mean(sample_sums(ps_clean_filter))
      range(sample_sums(ps_clean_filter))
```      
      

## Sample Rarefaction 

```{r}

# #GG Version Using MicrobiotaProcess
#     set.seed(1024) # for reproducibly random number
#     options(warn=-1)
#     rareres <- get_rarecurve(obj=ps_clean_filter,chunks=100)
# 
#   #Plot  
#     p_rare <- ggrarecurve(obj=rareres,
#                       indexNames=c("Observe","Shannon"),
#                       ) +
#           theme(legend.spacing.y=unit(0.01,"cm"),
#                 legend.text=element_text(size=4))
# 
# taxa_are_rows(ps_clean_filter)
# Base R/ Vegan Implementation 
  badger_matrix<-  t(as(otu_table(ps_clean_filter),"matrix"))
  raremax <- min(rowSums(badger_matrix))

  rarecurve(badger_matrix,step=50,cex=0.5,sample=raremax)


```

## Final Processing and Sample Summary Stats

### Sample Stats

```{r}

#Badgers
  length(unique(sample_data(ps_clean_filter)$id))


#Infection
    table(sample_data(ps_clean_filter)$day0_cdp)

#Known Age
    table(sample_data(ps_clean_filter)$age_fc) 
    length(unique(sample_data(ps_clean_filter)$id[sample_data(ps_clean_filter)$age_fc =="CUB"]))

```

# More Metadata

## Infection
```{r}

#Badgers with Longitudinal Test Data
  infection_meta<-read.csv('Badger Tb Test Data for Spatial.csv')

#Infection Grouping
  infection_meta$btb_status<-"Unexposed"
  infection_meta$btb_status[which(infection_meta$culture=="Positive")]<-"Excreter"
  infection_meta$btb_status[which(infection_meta$igy=="Positive" & infection_meta$culture=="Negative"|infection_meta$statpak2=="Positive" & infection_meta$culture=="Negative")]<-"Exposed"
  table(infection_meta$btb_status)
  
#Carry Across Swab Date
  woodchester_meta<-as(sample_data(ps_clean_filter),"data.frame")
  
#Test Left Join
  woodchester_meta_btb<-left_join(woodchester_meta,infection_meta,"id")
  
## DId Culture Test Occur Before Microbiome Swab
    woodchester_meta_btb$culture_before_sample<-with(woodchester_meta_btb,culture_first_pos <= sampledate_julian)
  ## Did IG Test Occur Before Microbiome Swab
    woodchester_meta_btb$igy_before_sample<-with(woodchester_meta_btb,ig_first_pos <= sampledate_julian) 
  ## Did StatPak Test Occur Before Microbiome Swab
    woodchester_meta_btb$statpak_before_sample<-with(woodchester_meta_btb,statpak_first_pos <= sampledate_julian)
    
  ### Tabulate To See Variation 
    with(woodchester_meta_btb,table(culture_before_sample,id))
    
  ## Create New Variable
      btb_status_time<-rep(NA,nrow(woodchester_meta_btb))
      btb_status_time[which(woodchester_meta_btb$culture=="Positive" & woodchester_meta_btb$culture_before_sample==FALSE)]<-"Future Excreter" #flag as these might have been exposed 
      btb_status_time[which(woodchester_meta_btb$culture=="Positive" & woodchester_meta_btb$culture_before_sample==TRUE)]<-"Excreter"
      btb_status_time[which(woodchester_meta_btb$btb_status=="Unexposed")]<-"Unexposed"
      btb_status_time[which(woodchester_meta_btb$igy=="Positive" & woodchester_meta_btb$igy_before_sample==TRUE  & woodchester_meta_btb$culture=="Negative")]<-"Exposed"
        btb_status_time[which(woodchester_meta_btb$statpak2=="Positive" & woodchester_meta_btb$statpak_before_sample==TRUE & woodchester_meta_btb$culture=="Negative")]<-"Exposed"
      btb_status_time[which(woodchester_meta_btb$igy=="Positive" & woodchester_meta_btb$igy_before_sample==FALSE  & woodchester_meta_btb$culture=="Negative" & woodchester_meta_btb$statpak2!="Positive")]<-"Future Exposed"
       btb_status_time[which(woodchester_meta_btb$statpak2=="Positive" & woodchester_meta_btb$statpak_before_sample==FALSE  & woodchester_meta_btb$culture=="Negative" & woodchester_meta_btb$igy!="Positive")]<-"Future Exposed"
   

#Now Sort Out the Future Excreters - Were these exposed?
    woodchester_meta_btb_fe<- subset(woodchester_meta_btb,btb_status_time=="Future Excreter")
    woodchester_meta_btb_fe$ig_first_pos< woodchester_meta_btb_fe$sampledate_julian #all exposed 
    
    if(all(woodchester_meta_btb_fe$ig_first_pos< woodchester_meta_btb_fe$sampledate_julian)){
      print("All Future Excreters Have a Positive IG result")
    } else {stop("Not all badgers exposed before test result")}
    
    
### Create Collapsed Version 
    btb_status_time_collapsed<-btb_status_time
    btb_status_time_collapsed[which(btb_status_time_collapsed=="Future Excreter")]<-"Exposed"
        btb_status_time_collapsed[which(btb_status_time_collapsed=="Future Exposed")]<-"Unexposed"

    ### Tabulate Overall
     table(btb_status_time_collapsed,useNA="always")
     
    ### Tabulate bu ID
        table(btb_status_time,woodchester_meta_btb$id)
      
### Add on to DF
  woodchester_meta_btb$btb_status_time<-btb_status_time_collapsed


### Variation in Time Since Excreting?
      woodchester_meta_btb_excreter<-woodchester_meta_btb %>% filter(btb_status_time=="Excreter")   
     range(woodchester_meta_btb_excreter$sampledate_julian -  woodchester_meta_btb_excreter$culture_first_pos,na.rm=T)

### Variation in Time Since Exposed?
     woodchester_meta_btb_exposed<-woodchester_meta_btb %>% filter(btb_status_time=="Exposed")   
     range(woodchester_meta_btb_exposed$sampledate_julian -  woodchester_meta_btb_exposed$ig_first_pos,na.rm=T)
      range(woodchester_meta_btb_exposed$sampledate_julian -  woodchester_meta_btb_exposed$statpak_first_pos,na.rm=T)     

#### REBIND ONTO PHYLOSEQ
      rownames(woodchester_meta_btb)<-woodchester_meta_btb$Sample
      sample_data(ps_clean_filter)<-woodchester_meta_btb

```

## Comparison of Continuous Tb Infection vs Categorical
```{r}
exposed_day0_comparisonplot1<-woodchester_meta_btb %>% filter(!is.na(btb_status_time)) %>% ggplot(.,aes(x=btb_status_time,y=day0_cdp))  + geom_point(position = position_jitter(width=0.1),shape=21,size=5,fill="white") + theme_bw(base_size=15) + labs(x="Categorical Infection Status",y="Infection Probability")
ggsave2('Categorical Tb Exposure Comparison Plot.png',exposed_day0_comparisonplot1)

#subset(woodchester_meta_btb,btb_status_time=="Unexposed" & day0_cdp>0.25)
#subset(woodchester_meta_btb,btb_status_time=="Exposed" & day0_cdp>0.75)

```

## Global Tb Category Colours
```{r}
## Colours 

 tbcols<-moma.colors("Palermo",n=3)
  names(tbcols)<-c("Unexposed","Excreter","Exposed")
```

# Alpha Diversity 

## Rarefy
```{r}

#Minimum Read Depth   
  min(sample_sums(ps_clean_filter))

#Rarefy 
  ps_rare<-rarefy_even_depth(ps_clean_filter,rngseed = 10112022,sample.size = 13569)
 

#Richness
  badger_rich<-estimate_richness(ps_rare,measures = c("Observed","Shannon"))
  badger_rich$Sample<- sample_data(ps_rare)$Sample
  badger_rich<-left_join(badger_rich,woodchester_meta_btb,"Sample")
  
##Filter
  table(badger_rich$social_group,useNA="ifany")
  badger_rich$year_social<-with(badger_rich,paste(capture_year,social_group,sep="_"))
  year_social_table<-table( badger_rich$year_social)
  
#Filter to Only Cases Where there are 3 or more data points
  year_social3<-names(year_social_table)[year_social_table>2]
  badger_rich_filter3<-subset(badger_rich,year_social %in% year_social3)

```

## Richness Plots

### Social Group
```{r}

## All Badgers
  richplot1<-ggplot(badger_rich,aes(x=social_group,y=Observed)) + geom_point(fill=moma.colors("Levine2")[3],colour="white",shape=21,size=5)+ facet_grid(capture_year~.)
  richplot2<- richplot1 + stat_summary(fun.data = "mean_cl_boot", colour = "black", linewidth = 1,shape=23,fill="white") + theme_bw() + labs(x="Social Group",y="Observed Bacterial \n Richness") + guides(fill="none") + theme(axis.text.x = element_text(angle = 45, hjust=1)) + theme(strip.text.y = element_text(size=18),axis.text = element_text(size=12),axis.title=element_text(size=15))
  richplot2
    #ggsave2('Richness All Badgers All Years.pdf',richplot2,width=15,height=15,units="cm")
      #ggsave2('Richness All Badgers All Years.tiff',richplot2,width=15,height=15,units="cm")


#Tabulate
    badger_social_tab<-with(badger_rich,table(social_group,capture_year))  
    #write.csv(badger_social_tab,'Badger Social Group By Year.csv')
      
      
```

### bTb Status - Categories 
```{r}
badger_rich$btb_status_time<-factor(badger_rich$btb_status_time,levels=c("Unexposed","Exposed","Excreter"))

 richplot_tbcat1<-filter(badger_rich,!is.na(btb_status_time)) %>% ggplot(.,aes(x=btb_status_time,y=Observed)) + geom_boxplot(outliers = FALSE) + geom_point(size=5,shape=21,colour="white",aes(fill=btb_status_time),position = position_jitter(width=0.2),alpha=0.7) + theme_bw(base_size = 15)
 richplot_tbcat2<- richplot_tbcat1 + labs(fill="bTb Status",y="Observed Bacterial Richness",x="bTb Status") + scale_fill_manual(values = tbcols) + guides(fill="none")
 richplot_tbcat2
 ggsave2('Richness by bTb Category.png',richplot_tbcat2)
```

### Richness by bTb Category by Individual 
```{r}
richplot_tbcat2 + facet_wrap(.~id)
```

### bTb Status - Probability of Infection 
```{r}

tb_lab<-expression(paste(italic("M. bovis  "), "Infection Probability"))

#Sample size
  nrow(badger_rich)
  
#All Badgers
  richplot_tb1<-ggplot(badger_rich,aes(x=day0_cdp,y=Observed)) + geom_smooth(method="lm") + geom_point(size=5,shape=21,colour="white",fill=moma.colors("Levine2")[3]) + facet_grid(.~capture_year) + theme_bw(base_size = 15)
 richplot_tb2<- richplot_tb1 + labs(x=tb_lab,y="Observed Bacterial \n Richness") 
 richplot_tb2
  #ggsave2('All Badgers Continuous Tb Status by Year.pdf',richplot_tb2,width=22,height=10,units="cm")

```

### Tb Within Social Group - Continuous

```{r}

##Filter To At Least 5 Per Group 
  rich_filter_tb_tab<-with(badger_rich,table(social_group))
  rich_filter_tb_tab_filtergroups<-rownames(rich_filter_tb_tab)[rich_filter_tb_tab>4]
  rich_filter_tb_filter<-subset(badger_rich,social_group %in% rich_filter_tb_tab_filtergroups)

#All Badgers
  richplot_tb_soc1<-ggplot(rich_filter_tb_filter,aes(x=day0_cdp,y=Observed)) + geom_smooth(method="lm") + geom_point(size=5,shape=21,colour="white",fill=moma.colors("Levine2")[3]) + facet_wrap(.~social_group) + theme_bw(base_size = 15) + labs(x=tb_lab,y="Observed Bacterial \n Richness")
  richplot_tb_soc1
  #ggsave2('All Badgers Tb Status by Social Group.pdf',richplot_tb_soc1,width=30,height=22,units="cm")


```

### Tb Within Social Group - Category

```{r}

##Filter To At Least 5 Per Group 
  rich_filter_tb_tab<-with(badger_rich,table(social_group))
  rich_filter_tb_tab_filtergroups<-rownames(rich_filter_tb_tab)[rich_filter_tb_tab>4]
  rich_filter_tb_filter<-subset(badger_rich,social_group %in% rich_filter_tb_tab_filtergroups & !is.na(btb_status_time))

#All Badgers
  richplot_tb_statustime_soc1<-ggplot(rich_filter_tb_filter,aes(x=btb_status_time,y=Observed)) + geom_smooth(method="lm") + geom_point(size=5,shape=21,colour="white",aes(fill=btb_status_time),position = position_jitter(width=0.2)) + facet_wrap(.~social_group) + theme_bw(base_size = 15) + labs(x="bTb Status",y="Observed Bacterial \n Richness",fill="bTb Status") + scale_fill_manual(values=tbcols) + theme(axis.text.x=element_text(angle=45,hjust=1),legend.position="top")
  richplot_tb_statustime_soc1
  #ggsave2('All Badgers Tb Categorical Status by Social Group.png',richplot_tb_statustime_soc1,width=30,height=22,units="cm")


```

### Richness in Excreters as a Function of Time Since Infection 
```{r}
badger_rich_excreter<-subset(badger_rich,btb_status_time=="Excreter")
badger_rich_excreter$delta_time<-with(badger_rich_excreter,sampledate_julian - culture_first_pos)

ggplot(badger_rich_excreter,aes(x=delta_time,y=Observed)) + geom_point(shape=21,size=5,aes(fill=id)) + theme_bw(base_size = 15)
```



### Tb Combined Plot
```{r}
tb_richness_plot1<-plot_grid(richplot_tb_soc1,richplot_tb2,nrow=2,labels="AUTO",label_size = 22,rel_heights = c(2,1))
tb_richness_plot1
#ggsave2('All Badgers Richness Continuous Tb Combined Plot.pdf',tb_richness_plot1,width=20,height=22,units="cm")
#ggsave2('Richness Tb Combined Plot.tiff',tb_richness_plot1,width=20,height=22,units="cm")

```

### Mass & Condition Score 

```{r}

## Relevel COndition
    badger_rich$condition<-factor(badger_rich$condition,levels=c("POOR","FAIR","GOOD","VERY GOOD"))
  
            #Richness Plot  by MASS 
              mass_plot1<-badger_rich %>% filter(!is.na(weight)) %>% ggplot(.,aes(x=weight,y=Observed))  + geom_smooth(method="lm")  + geom_point(shape=21,aes(fill=condition),size=5,color="white",alpha=0.7) + theme_bw()
              mass_plot2 <- mass_plot1+ theme(axis.text=element_text(size=20),axis.title = element_text(size=20),strip.text.x=element_text(size=20)) + labs(y="Observed Bacterial \n Richness",x="Mass (kg)")
              mass_plot3<-mass_plot2   + facet_wrap(capture_year~.)  + labs(fill="Condition") + scale_fill_brewer(palette = "Paired")
              mass_plot3
             # ggsave2('Mass by Year.pdf',mass_plot3,width=25,height=10,units="cm")
            

```

## Richness Models 

### Data Prep
```{r}
library(brms)

## Sex Variable
  #badger_rich$sex<-ifelse(is.na(badger_rich$testes),"F","M")
  with(badger_rich,table(reproc,testes,exclude = NULL))
  subset(badger_rich,is.na(reproc) & is.na(testes))
  
##Filter Data to No Missing Values for Any Predictor   
  badger_rich_complete<-badger_rich[with(badger_rich,complete.cases(weight,capture_year,toothwear,social_group,day0_cdp)),]
  
##Summary Stats  
  nrow(badger_rich_complete) # 165 samples
  length(unique(badger_rich_complete$social_group)) #18 social groups (18 in BRMS model)
  length(unique(badger_rich_complete$capture_year)) #3 capture years
  length(unique(badger_rich_complete$id)) #72 badgers

##Factor for Cap Year  
  badger_rich_complete$capture_year<-factor(badger_rich_complete$capture_year)

```

### Plots
```{r}
## Check Correlation Between Age and Mass
    ggplot(badger_rich_complete,aes(x=known_age,y=weight)) + geom_smooth(method="lm",formula=y~x+I(x^2)) + geom_point(shape=21,size=5,fill="white")
    with(badger_rich_complete,cor.test(known_age,weight))
    
  ### All Data Plot  
   ggplot(badger_rich_complete,aes(x=known_age,y=Observed)) + geom_smooth(method="lm") + geom_point(shape=21,size=5,fill="white") + scale_x_continuous(n.breaks=11) + theme_bw() + plotopts + labs(y="Observed Richness",x="Age")
    with(badger_rich_complete,cor.test(known_age,Observed))
    
  ### Subset to Repeat Sampled Badgers
    badger_rich_tab<-table(badger_rich_complete$id,badger_rich_complete$known_age)
    badger_rich_tab2<-apply(badger_rich_tab,1,function(x)sum(x>0))
    badger_rich_complete_min3<-subset(badger_rich_complete,id %in% names(badger_rich_tab2)[badger_rich_tab2>2])
    table(badger_rich_complete_min3$id)
    
#### PLOT    
   ggplot(badger_rich_complete_min3,aes(x=known_age,y=Observed,fill=id)) + geom_smooth(method="lm",se=F) + geom_point(shape=21,size=5) + theme_bw() + plotopts + labs(y="Observed Richness",x="Age") + scale_x_continuous(n.breaks=11) + facet_wrap(.~id) + guides(fill="none")
   
```

### Models

#### Prepare Data
```{r}
#Fix Capture Year 
      badger_rich_complete$capture_year<-factor(badger_rich_complete$capture_year)
      badger_rich_complete$toothwear<-as.factor(badger_rich_complete$toothwear)

#Sex
    subset(badger_rich_complete,is.na(reproc) & is.na(testes))  
    badger_rich_complete$sex<-ifelse(is.na(badger_rich_complete$testes),"F","M")
    badger_rich_complete$sex[badger_rich_complete$id=="ID27"]<-"M"
    badger_rich_complete$sex[badger_rich_complete$id=="ID32"]<-"F"
    
    
## Subset to Known Age
    badger_rich_complete_age<-subset(badger_rich_complete,!is.na(known_age))
    
#Sample Size
    nrow(badger_rich_complete_age)
    length(unique(badger_rich_complete_age$id))
    length(unique(badger_rich_complete_age$social_group))

```

## Bivariate Model - Continuous bTb Infection
```{r}

library(brms)

  
    #Model For Each Outcome 
        bf_rich<-bf(Observed ~ capture_year +  known_age + I(known_age^2) + sex + day0_cdp +  (1|p|social_group) + (1|q|id)) + poisson()
        bf_mass<-bf(weight ~ capture_year  + known_age + I(known_age^2)  + sex + day0_cdp + (1|p|social_group) + (1|q|id)) + gaussian()

  fit1 <- brm(
  bf_rich + bf_mass + set_rescor(FALSE),
  data = badger_rich_complete_age, chains = 4, cores = 4,
  control = list(adapt_delta = 0.95),save_pars = save_pars(all = TRUE),
)

  summary(fit1)
  conditional_effects(fit1)
  bayes_R2(fit1)
  
  
  pp_check(fit1,resp='Observed')
  pp_check(fit1,resp='weight')

  
######## NULL Model Checking 
      #Model For Each Outcome 
        bf_rich_null<-bf(Observed ~ 1+  (1|p|social_group) + (1|q|id)) + poisson()
        bf_mass_null<-bf(weight ~ 1 + (1|p|social_group) + (1|q|id)) + gaussian()
  ##Fit NULL      
    fit_null <- brm(
    bf_rich_null + bf_mass_null + set_rescor(FALSE),
    data = badger_rich_complete_age, chains = 4, cores = 4,
    control = list(adapt_delta = 0.95),save_pars = save_pars(all = TRUE),
      )
    summary(fit_null)
  #Add Criteria  
        fit_null<-add_criterion(fit_null,"kfold",K=10)
        fit1<-add_criterion(fit1,"kfold",K=10)

    print(loo_compare(fit_null,fit1,criterion = "kfold"),simplify = F)
    
####### EPLD Checking
    
```

### Model Plots
```{r}
## Richness With Age
  agerichdata<- conditional_effects(fit1)[[3]]
  agerichplot<-ggplot() + geom_ribbon(data=agerichdata,aes(x=known_age,ymax=upper__,ymin=lower__),fill="lightgray") + geom_line(data=agerichdata,aes(x=known_age,y=estimate__),lwd=1.5) + theme_bw(base_size = 15) + geom_point(data=badger_rich_complete_age,aes(x=known_age,y=Observed,fill=day0_cdp),shape=21,size=5,color="white") + labs(x="Chronological Age",y="Bacterial Richness",fill="Infection Probability") + scale_fill_moma_c("Exter") + theme(legend.position = "top",legend.text = element_text(size=8))
  agerichplot
  
## MAss WIth Age 
   agemassdata<- conditional_effects(fit1,conditions = data.frame(sex=c("M","F")))[[4]]
   
  agemassplot<-ggplot() + geom_ribbon(data=agemassdata,aes(x=known_age,ymax=upper__,ymin=lower__,fill=sex),alpha=0.5) + geom_line(data=agemassdata,aes(x=known_age,y=estimate__,color=sex),lwd=1.5) + theme_bw(base_size = 15) + geom_point(data=badger_rich_complete_age,aes(x=known_age,y=weight,fill=sex),shape=21,size=5,color="white") + labs(x="Chronological Age",y="Mass (kg)",fill="Sex") + scale_fill_moma_d("Smith") + guides(color="none")+ theme(legend.position = "top")
  agemassplot
  
bivariateplot1<-plot_grid(agerichplot,agemassplot,nrow=1,labels="AUTO",label_size = 20) 
ggsave2('Age Mass Richness Bivariate Model Plot.png',bivariateplot1,width=30,height=20,units="cm")
```

## Univariate Response Richness by Btb Category 
```{r}

#Subset to Key Categories
  badger_rich_tbcat<-subset(badger_rich_complete,!is.na(btb_status_time))
  table(badger_rich_tbcat$btb_status_time)
  
#Model 
   badger_richmod_tbcat1<-brm(Observed ~ capture_year +  known_age + I(known_age^2) + sex + btb_status_time +  (1|p|social_group) + (1|q|id),data=badger_rich_tbcat, family=negbinomial())
   summary(badger_richmod_tbcat1)
```

## Bivariate Model - Categorical bTb Infection
```{r}

library(brms)

  
    #Model For Each Outcome 
        bf_rich_tbstatus<-bf(Observed ~ capture_year +  known_age + I(known_age^2) + sex + btb_status_time +  (1|p|social_group) + (1|q|id)) + poisson()
        bf_mass_tbstatus<-bf(weight ~ capture_year  + known_age + I(known_age^2)  + sex + btb_status_time + (1|p|social_group) + (1|q|id)) + gaussian()

  fit_tbstatus1 <- brm(
  bf_rich_tbstatus + bf_mass_tbstatus + set_rescor(FALSE),
  data = badger_rich_tbcat, chains = 4, cores = 4,
  control = list(adapt_delta = 0.95),save_pars = save_pars(all = TRUE),
)

  summary(fit_tbstatus1)
  conditional_effects(fit_tbstatus1)
  
  bayes_R2(fit_tbstatus1)

  
  pp_check(fit_tbstatus1,resp='Observed')
  pp_check(fit_tbstatus1,resp='weight')


```


# Beta Diversity

## Some Functions

```{r}

#Sample for Stripping out the ASV matrix from a phyloseq object (run all of this)
  vegan_otu <- function(physeq) {
    OTU <- otu_table(physeq)
    if (taxa_are_rows(OTU)) {
      OTU <- t(OTU)
    }
    return(as(OTU, "matrix"))
  }
  

```

## Stacked Barplots

```{r}

## Top 20 Families 
  badger_topfam <- aggregate_top_taxa2(ps_clean_filter, "Family", top = 20)

#Plot Sample Level 
  badger_topfam %>%  
  microbiome::transform(transform = "compositional") %>%
  plot_composition()

```


### Year Grouping
```{r}

#Extract and Reorder
  fams<-tax_table(badger_topfam)[,2] %>% as.vector()
  fams_ordered<-c("Other",fams[-which(fams=="Other")])     

barpal<-kelly.colors(n=21); names(barpal)<-fams

#Remake Our Graph  but with averaging by YEAR
  physeq_topfam_year_plot <-  badger_topfam %>%  
  microbiome::transform(transform = "compositional") %>%
  plot_composition(average_by = "capture_year") + scale_fill_manual(values = barpal) + theme_classic() + labs(fill="Bacterial Family",x="Capture Year") + plotopts2 + theme(axis.text.x = element_text(angle=45,hjust=1, size = 14), axis.text.y = element_text(size = 14)) 

  physeq_topfam_year_plot
    #ggsave2('Badger Barplot by Year.pdf',physeq_topfam_year_plot,width=10,height=15,units="cm")
    #ggsave2('Badger Barplot by Year.tiff',physeq_topfam_year_plot,width=10,height=15,units="cm")
```

### Social Group Grouping
```{r}
    #Remake Our Graph  but with averaging by Social Group
  physeq_topfam_social_plot <-  badger_topfam %>%  
  microbiome::transform(transform = "compositional") %>%
  plot_composition(average_by = "social_group") + scale_fill_manual(values = barpal) + theme_classic() + labs(fill="Bacterial Family",x="Social Group") + plotopts2 + theme(axis.text.x = element_text(angle=45,hjust=1, size = 14), axis.text.y = element_text(size = 14)) + guides(fill="none")

  physeq_topfam_social_plot
        #ggsave2('Badger Barplot by SocialGroup.pdf',physeq_topfam_social_plot,width=15,height=10,units="cm")
        #ggsave2('Badger Barplot by SocialGroup.tiff',physeq_topfam_social_plot,width=20,height=15,units="cm")

```

### Combined Plot 
```{r}
#Combined plot
abundance_plots<-plot_grid(physeq_topfam_social_plot,physeq_topfam_year_plot,nrow=2,labels=c("A", "B"), label_size = 25)
abundance_plots
ggsave("Combined_Badger_Barplots.png", abundance_plots, width = 20, height = 30, units = "cm")
#ggsave("Combined_Badger_Barplots.pdf", abundance_plots, width = 30, height = 20, units = "cm")


```

## CLR Transformation

#### Function to Extract Variance Explained Percentages
```{r}
  pcsum_func<-function(ordobject){
      pcsum<-summary(ordobject)
      pc1r2<-signif(pcsum$cont$importance[2, "PC1"]*100,digits=3) 
      pc2r2<-signif(pcsum$cont$importance[2, "PC2"]*100,digits=3) 
      return(c(pc1r2,pc2r2))
  }

```

### Prepare Data - Social Group Effects by Year
```{r}

#Subset All Samples
  ps_clean_filter_s<-as(sample_data(ps_clean_filter),"data.frame")

##Tabulate By Group and Year
    grouptab<-with(ps_clean_filter_s,table(social_group,capture_year))
  # 2016 Groups with 5 or more
    bagdergroups2016<-rownames(grouptab)[grouptab[,1]>3]
    bagdergroups2017<-rownames(grouptab)[grouptab[,2]>3]
  #Subset Per Year 2016
    ps_badger2016_orddata<-prune_samples(sample_data(ps_clean_filter)$social_group %in% bagdergroups2016 & sample_data(ps_clean_filter)$capture_year=="2016",ps_clean_filter) 
   #Subset Per Year 2017
    ps_badger2017_orddata<-prune_samples(sample_data(ps_clean_filter)$social_group %in% bagdergroups2017 & sample_data(ps_clean_filter)$capture_year=="2017",ps_clean_filter) 
   
     
#### CLR Transform
  ## Extract for Ordination and Plotting 2016  
    ps_orddata2016_clr<-microbiome::transform(ps_badger2016_orddata,"clr")
    badger_2016_clr_v<-vegan_otu(ps_orddata2016_clr)
    badger_2016_clr_s<-as(sample_data(ps_orddata2016_clr),"data.frame")
 ## Extract for Ordination and Plotting 2017  
    ps_orddata2017_clr<-microbiome::transform(ps_badger2017_orddata,"clr")
    badger_2017_clr_v<-vegan_otu(ps_orddata2017_clr)
    badger_2017_clr_s<-as(sample_data(ps_orddata2017_clr),"data.frame")
    
##### Sort Out Consistent Colours for Each Group  
    groups_bothyears<- c(bagdergroups2016,bagdergroups2017) %>% unique()
    groupcols<-moma.colors("Warhol")[1:length(groups_bothyears)]
    names(groupcols)<-groups_bothyears
```

### Social Group PCA Ordinations 2016
```{r}
    #Run PCA 
      woodchester_2016_clr_pca<-rda(badger_2016_clr_v)

    #Extract Variance Explained
      social2016_pcvariance<-pcsum_func(woodchester_2016_clr_pca)
  
  #Inspect Plot 
    gg_ordiplot(woodchester_2016_clr_pca,groups=badger_2016_clr_s$social_group,spiders = TRUE)
    
    #Extract for GGPLot2
      social2016_ggord_data<-gg_ordiplot(woodchester_2016_clr_pca,groups=badger_2016_clr_s$social_group,spiders = TRUE,plot=FALSE) 
      social2016_ggord_data2<-social2016_ggord_data$df_ord
      #social2016_ggord_data2$capture_year<-woodchester_group5_clr_s$capture_year
      
  ##### PERMANOVA
  badger_how_2016 <- how(nperm=9999, plots = Plots(strata=badger_2016_clr_s$id))
  adonis2(badger_2016_clr_v ~ social_group,data=badger_2016_clr_s,by="mar",method="euclidean",how=badger_how_2016)
  
  #### DISPERSION
   badger_2016_dispersion<-betadisper(vegdist(badger_2016_clr_v,"euclidean"),badger_2016_clr_s$social_group)
    permutest(badger_2016_dispersion) 
    
  #Label for Plot from PERMANOVA    
      badger2016graphlab<-paste("Social~Group~r^2==0.22\n~P==0.001")
  #Plot   
    badger_social2016_clr_plot1<-ggplot() + geom_segment(data=social2016_ggord_data$df_spiders,linewidth=1, aes(x=cntr.x, xend=x, y=cntr.y, yend=y, color=Group), show.legend = FALSE) 
  #Plot Points  
    badger_social2016_clr_plot2<- badger_social2016_clr_plot1 + geom_point(data=social2016_ggord_data2,aes(x=x,y=y,fill=Group),shape=21,color="white",size=5,alpha=0.8) + theme_bw() 
 #Colour
    badger_social2016_clr_plot3<- badger_social2016_clr_plot2 + labs(x=paste("PC1",social2016_pcvariance[1],"%"),y=paste("PC2",social2016_pcvariance[2],"%"),fill="Social Group") + plotopts + theme(legend.text = element_text(size=12), axis.text.y = element_text(size = 14), axis.text.x = element_text(size = 14)) + labs(fill="Social Group") + scale_fill_manual(values = groupcols) + scale_color_manual(values = groupcols)
#Add Ellipses
    badger_social2016_clr_plot4<- badger_social2016_clr_plot3 + geom_path(data = social2016_ggord_data$df_ellipse,linewidth=1, aes(x = x, y = y, color = Group), show.legend = FALSE) + annotate("text", x=-0.8, y=1.5, label=as.character(badger2016graphlab),size=6,parse=T) 
  badger_social2016_clr_plot4
    #ggsave2('CLR Beta Diversity Group5.pdf',badger_social1617_clr_plot4,width=25,height=10,units="cm")
    #ggsave2('CLR Beta Diversity Group5.png',badger_social1617_clr_plot4,width=25,height=15,units="cm")
    



```

### Social Group PCA Ordinations 2017
```{r}
    #Run PCA 
      woodchester_2017_clr_pca<-rda(badger_2017_clr_v)

    #Extract Variance Explained
      social2017_pcvariance<-pcsum_func(woodchester_2017_clr_pca)
  
  #Inspect Plot 
    gg_ordiplot(woodchester_2017_clr_pca,groups=badger_2017_clr_s$social_group,spiders = TRUE)
    
    #Extract for GGPLot2
      social2017_ggord_data<-gg_ordiplot(woodchester_2017_clr_pca,groups=badger_2017_clr_s$social_group,spiders = TRUE,plot=FALSE) 
      social2017_ggord_data2<-social2017_ggord_data$df_ord
      #social2016_ggord_data2$capture_year<-woodchester_group5_clr_s$capture_year
      
##### PERMANOVA
  badger_how_2017 <- how(nperm=9999, plots = Plots(strata=badger_2017_clr_s$id))
  adonis2(badger_2017_clr_v ~ social_group,data=badger_2017_clr_s,by="mar",method="euclidean",how=badger_how_2017)
  
#### DISPERSION
   badger_2017_dispersion<-betadisper(vegdist(badger_2017_clr_v,"euclidean"),badger_2017_clr_s$social_group)
    permutest(badger_2017_dispersion)
    
  #Label for Plot from PERMANOVA    
      badger2017graphlab<-paste("Social~Group~r^2==0.16\n~P==0.003")
  #Plot   
    badger_social2017_clr_plot1<-ggplot() + geom_segment(data=social2017_ggord_data$df_spiders,linewidth=1, aes(x=cntr.x, xend=x, y=cntr.y, yend=y, color=Group), show.legend = FALSE) 
  #Plot Points  
    badger_social2017_clr_plot2<- badger_social2017_clr_plot1 + geom_point(data=social2017_ggord_data2,aes(x=x,y=y,fill=Group),shape=21,color="white",size=5,alpha=0.8) + theme_bw() 
 #Colour
    badger_social2017_clr_plot3<- badger_social2017_clr_plot2 + labs(x=paste("PC1",social2017_pcvariance[1],"%"),y=paste("PC2",social2017_pcvariance[2],"%"),fill="Social Group") + plotopts + theme(legend.text = element_text(size=12), axis.text.y = element_text(size = 14), axis.text.x = element_text(size = 14)) + labs(fill="Social Group") + scale_fill_manual(values = groupcols) + scale_color_manual(values = groupcols)
#Add Ellipses
    badger_social2017_clr_plot4<- badger_social2017_clr_plot3 + geom_path(data = social2017_ggord_data$df_ellipse,linewidth=1, aes(x = x, y = y, color = Group), show.legend = FALSE) + annotate("text", x=0.8, y=1.5, label=as.character(badger2017graphlab),size=6,parse=T) 
  badger_social2017_clr_plot4
```

### Prepare Data All Samples Complete Cases
```{r}

#Extract Matrix and Sample Data - Same Dataset for Richness
  ps_badger_forclr<-prune_samples(sample_data(ps_clean_filter)$Sample %in% badger_rich_complete_age$Sample,ps_clean_filter)

#CLR Transform 
  ps_clr<-microbiome::transform(ps_badger_forclr,"clr")
  woodchester_clr_v<-vegan_otu(ps_clr)
  woodchester_clr_s<-as(sample_data(ps_clr),"data.frame")
  woodchester_clr_s$sex<-badger_rich_complete$sex[match(woodchester_clr_s$Sample,badger_rich_complete$Sample)]
  
```

### Prepare Data Tb Category Exposed/Unexposed/Excreter
```{r}

#Extract Matrix and Sample Data - Same Dataset for Richness
  ps_badger_tb3cat<-prune_samples(sample_data(ps_clean_filter)$Sample %in% badger_rich_tbcat$Sample,ps_clean_filter)

#CLR Transform 
  ps_tb3cat_clr<-microbiome::transform(ps_badger_tb3cat,"clr")
  woodchester_tb3cat_clr_v<-vegan_otu(ps_tb3cat_clr)
  woodchester_tb3cat_clr_s<-as(sample_data(ps_tb3cat_clr),"data.frame")
  woodchester_tb3cat_clr_s$sex<-badger_rich_complete$sex[match(woodchester_tb3cat_clr_s$Sample,badger_rich_complete$Sample)]
  
  
### ORDINATE
  woodchester_tb3cat_clr_pca<-rda(woodchester_tb3cat_clr_v)

#Extract Variance Explained
  woodchester_tb3cat_pcvariance<-pcsum_func(woodchester_tb3cat_clr_pca)
  
  #Inspect Plot 
    gg_ordiplot(woodchester_tb3cat_clr_pca,groups=woodchester_tb3cat_clr_s$btb_status_time,spiders = TRUE)

```



## PERMANOVA

### Categorical TB 
```{r}
    
  #Betadisper 
    badger_tb3cat_d<-vegdist(woodchester_tb3cat_clr_v,"euclidean")
    badger_tb3cat_dispersion<-betadisper(badger_tb3cat_d,woodchester_tb3cat_clr_s$btb_status_time)
    permutest(badger_tb3cat_dispersion)
    
    
  #PERMANOVA
    badger_how_cattb <- with(woodchester_tb3cat_clr_s,how(nperm=9999, blocks=id))

    badger_tb3cat_clr_perm<-adonis2(woodchester_tb3cat_clr_v ~ id + btb_status_time + factor(capture_year) +factor(social_group) ,data=woodchester_tb3cat_clr_s,method="euclidean",by="mar",how=badger_how_cattb)
    badger_tb3cat_clr_perm
    
  
#   #Individual ID     
#   badger_tb3cat_clr_id_perm<-adonis2(woodchester_tb3cat_clr_v ~ factor(id) ,data=woodchester_tb3cat_clr_s,method="euclidean",by="mar",how=badger_how_cattb)
# badger_tb3cat_clr_id_perm

```

### Continuous Tb 
```{r}
### Complete Cases
  nrow(woodchester_clr_s)
  length(unique(woodchester_clr_s$social_group))
  length(unique(woodchester_clr_s$id))
  table(woodchester_clr_s$capture_year)
  mean(complete.cases(woodchester_clr_s[,c("known_age","day0_cdp","social_group","sex","capture_year")])) #all complete data
    
  woodchester_clr_s$social_group<-factor(woodchester_clr_s$social_group)


### BETADISPER to Check Dispersion Assumptions

#Distance Matrix
    woodchester_clr_d<-vegdist(woodchester_clr_v,"euclidean")
#Social Group
    badger_social_dispersion<-betadisper(woodchester_clr_d,woodchester_clr_s$social_group)
    permutest(badger_social_dispersion)
#Capture Year
    badger_captureyear_dispersion<-betadisper(woodchester_clr_d,woodchester_clr_s$capture_year)
    permutest(badger_captureyear_dispersion)

### ADONIS / PERMANOVA

#Constraint of ID
    badger_how <- with(woodchester_clr_s,how(nperm=9999, blocks=id))

#Fit Model Type III SS
  badger_clr_perm<-adonis2(woodchester_clr_v ~  social_group + factor(capture_year) + known_age + day0_cdp  ,data=woodchester_clr_s,method="euclidean",by="mar",permutations=badger_how)
  badger_clr_perm
  
```


## Ordinations

### Prepare Data 
```{r}
  library(ggordiplots)
  
  ##Tabulate By Group and Year
    grouptab<-with(woodchester_clr_s,table(social_group,capture_year))
  # 2016 Groups with 5 or more
    bagdergroups2016<-rownames(grouptab)[grouptab[,1]>3]
    bagdergroups2017<-rownames(grouptab)[grouptab[,2]>3]
  #Subset Per Year 2016
    ps_badger2016_orddata<-prune_samples(sample_data(ps_badger_forclr)$social_group %in% bagdergroups2016 & sample_data(ps_badger_forclr)$capture_year=="2016",ps_badger_forclr) 
   #Subset Per Year 2017
    ps_badger2017_orddata<-prune_samples(sample_data(ps_badger_forclr)$social_group %in% bagdergroups2017 & sample_data(ps_badger_forclr)$capture_year=="2017",ps_badger_forclr) 
   
```



### PCA Ordination (CLR) Individual ID
```{r}
##Tabulate By Group and Year
    idtab<-with(sample_data(ps_clean_filter),table(id))
  # 3 or more Obs
    id3plus<-names(idtab)[idtab>2]
    ps_badger_id_orddata<-prune_samples(sample_data(ps_clean_filter)$id %in% id3plus,ps_clean_filter)
 
  ## Extract for Ordination and Plotting 2016  
    ps_badger_id_clr<-microbiome::transform(ps_badger_id_orddata,"clr")
    badger_id_clr_v<-vegan_otu(ps_badger_id_clr)
    badger_id_clr_s<-as(sample_data(ps_badger_id_clr),"data.frame")
    

    #Run PCA 
        id_clr_pca<-rda(badger_id_clr_v)
    #Variance
      id_variance<-pcsum_func(id_clr_pca)
  #Test Ordinate
        gg_ordiplot(id_clr_pca,groups=badger_id_clr_s$id,spiders = TRUE)

###### PLOT

      id_ggord_data<-gg_ordiplot(id_clr_pca,groups=badger_id_clr_s$id,spiders = TRUE,plot=FALSE) 
      id_ggord_data2<-id_ggord_data$df_ord

        #Plot   
          id_clr_plot1<-ggplot() + geom_segment(data=id_ggord_data$df_spiders,linewidth=1, aes(x=cntr.x, xend=x, y=cntr.y, yend=y, color=Group), show.legend = FALSE) 
        #Plot Points  
          id_clr_plot2<- id_clr_plot1 + geom_point(data=id_ggord_data2,aes(x=x,y=y,fill=Group),shape=21,color="white",size=5,alpha=0.8) + theme_bw() 
       #Colour
          id_clr_plot3<- id_clr_plot2 + labs(x=paste("PC1",id_variance[1],"%"),y=paste("PC2",id_variance[2],"%"),fill="ID") + plotopts + theme(legend.text = element_text(size=12), axis.text.y = element_text(size = 14), axis.text.x = element_text(size = 14)) 
      #Add Ellipses
          id_clr_plot4<- id_clr_plot3 + geom_path(data = id_ggord_data$df_ellipse,linewidth=1, aes(x = x, y = y, color = Group), show.legend = FALSE)
      id_clr_plot4
```

### Disease Class 
```{r}

 #Inspect Plot 
    gg_ordiplot(woodchester_tb3cat_clr_pca,groups=woodchester_tb3cat_clr_s$btb_status_time,spiders = TRUE)
 
  #Variance
      tb3cat_variance<-pcsum_func(woodchester_tb3cat_clr_pca)
  
 #Extract for GGPLot2
      tb3cat_ggord_data<-gg_ordiplot(woodchester_tb3cat_clr_pca,groups=woodchester_tb3cat_clr_s$btb_status_time,spiders = TRUE,plot=FALSE) 
      tb3cat_ggord_data2<-tb3cat_ggord_data$df_ord

  #Plot   
    tb3cat_clr_plot1<-ggplot() + geom_segment(data=tb3cat_ggord_data$df_spiders,linewidth=1, aes(x=cntr.x, xend=x, y=cntr.y, yend=y, color=Group), show.legend = FALSE) 
  #Plot Points  
    tb3cat_clr_plot2<- tb3cat_clr_plot1 + geom_point(data=tb3cat_ggord_data2,aes(x=x,y=y,fill=Group),shape=21,color="white",size=5,alpha=0.8) + theme_bw() 
 #Colour
    tb3cat_clr_plot3<- tb3cat_clr_plot2 + labs(x=paste("PC1",tb3cat_variance[1],"%"),y=paste("PC2",tb3cat_variance[2],"%"),fill="bTb Status") + plotopts + theme(legend.text = element_text(size=12), axis.text.y = element_text(size = 14), axis.text.x = element_text(size = 14)) + scale_fill_manual(values=tbcols) + scale_color_manual(values=tbcols)
#Add Ellipses
    tb3cat_clr_plot4<- tb3cat_clr_plot3 + geom_path(data = tb3cat_ggord_data$df_ellipse,linewidth=1, aes(x = x, y = y, color = Group), show.legend = FALSE) + theme(legend.position = "top")
  tb3cat_clr_plot4
```

### Combined Plot - Beta Diversity
```{r}

ordinationplot1<-plot_grid(badger_social2016_clr_plot4,badger_social2017_clr_plot4,nrow=2,labels="AUTO",label_size = 30)
ordinationplot1
ggsave2('Badger Ordination Grid.png',ordinationplot1,width=25,height=35,units="cm")

```

### COmbined Plot - Categorical Tb Alpha & Beta 
```{r}

cattb_plot1<-plot_grid(richplot_tbcat2,richplot_tb_statustime_soc1,tb3cat_clr_plot4,nrow=3,labels="AUTO",label_size = 20,rel_heights = c(1,1.5,1))
ggsave2('Categorical Tb Composite Plot.png',cattb_plot1,width=20,height=40,units="cm")
```


# Individual Dynamics 

## Extract Values
```{r}
   
   ### All Badger CLR 
    ps_all_clr<-microbiome::transform(ps_clean_filter,"clr")
    ps_all_clr_v<-vegan_otu(ps_all_clr)
    ps_all_clr_s<-as(sample_data(ps_all_clr),"data.frame")
    ps_all_ord<-rda(ps_all_clr_v)
    ps_all_scores<-scores(ps_all_ord)$sites
    ps_all_scores2<-cbind(ps_all_scores,ps_all_clr_s)
  
      ps_all_scores2$Sample<-rownames(ps_all_scores2)
  ps_all_scores2$observed_richness<-badger_rich$Observed[match(ps_all_scores2$Sample,badger_rich$Sample)]
#Subset
  ps_all_scores3<-subset(ps_all_scores2,!is.na(known_age))
#Sample Size  
  nrow(ps_all_scores3)
  length(unique(ps_all_scores3$id))


```


## Age PC1 Individual Level 
```{r}
  #### Dynamics with Age for Repeat Sampled Badger
    badger_beta_tab<-table(ps_all_scores2$id,ps_all_scores2$known_age)
    badger_beta_tab2<-apply(badger_beta_tab,1,function(x)sum(x>0))
    badger_beta_complete_min3<-subset(ps_all_scores2,id %in% names(badger_beta_tab2)[badger_beta_tab2>2])
    table(badger_beta_complete_min3$id)
    
## Factor 
      badger_beta_complete_min3$btb_status_time<-factor(badger_beta_complete_min3$btb_status_time,levels=c("Unexposed","Exposed","Excreter"))
      
#### PLOT    
   pc1_age_plot1<-ggplot(badger_beta_complete_min3,aes(x=known_age,y=PC1,fill=id)) + geom_smooth(method="lm",se=F) + geom_point(shape=21,size=5) + theme_bw() + plotopts + labs(y="Beta Diversity (PC1)",x="Age") + scale_x_continuous(n.breaks=11) + facet_wrap(.~id) + guides(fill="none") + scale_fill_manual(values=MoMAColors::moma.colors("Levine2",6)) + theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.text.y = element_text(size = 14))
   pc1_age_plot1
   
#### PLOT with DIsease Class
   pc1_age_dc_plot1<-ggplot(badger_beta_complete_min3,aes(x=known_age,y=PC1,fill=btb_status_time)) + geom_smooth(method="lm",se=F) + geom_point(shape=21,size=5,colour="white") + theme_bw() + plotopts + labs(y="Beta Diversity (PC1)",x="Age",fill="bTb Status") + scale_x_continuous(n.breaks=11) + facet_wrap(.~id)  + scale_fill_manual(values=tbcols) + theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.text.y = element_text(size = 14),legend.position = "right")
   pc1_age_dc_plot1
```

### Age Richness Individual Level 

```{r}

 #### Dynamics with Age for Repeat Sampled Badger
    badger_rich_tab<-table(badger_rich$id,badger_rich$known_age)
    badger_rich_tab2<-apply(badger_rich_tab,1,function(x)sum(x>0))
    badger_rich_complete_min3<-subset(badger_rich,id %in% names(badger_rich_tab2)[badger_rich_tab2>2])
    table(badger_rich_complete_min3$id)
    
  badger_rich_complete_min3$btb_status_time<-factor(badger_rich_complete_min3$btb_status_time,levels=c("Unexposed","Exposed","Excreter"))
  
#### PLOT    
   rich_age_plot1<-ggplot(badger_rich_complete_min3,aes(x=known_age,y=Observed,fill=id)) + geom_smooth(method="lm",se=F) + geom_point(shape=21,size=5) + theme_bw() + plotopts + labs(y="Alpha Diversity (Richness)",x="Age") + scale_x_continuous(n.breaks=11) + facet_wrap(.~id) + guides(fill="none") + scale_fill_manual(values=MoMAColors::moma.colors("Levine2",6))+ theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.text.y = element_text(size = 14))
   rich_age_plot1
  
   rich_age_dc_plot1<-ggplot(badger_rich_complete_min3,aes(x=known_age,y=Observed,fill=btb_status_time)) + geom_smooth(method="lm",se=F) + geom_point(shape=21,size=5,colour="white") + theme_bw() + plotopts + labs(y="Alpha Diversity (Richness)",x="Age",fill="bTb Status") + scale_x_continuous(n.breaks=11) + facet_wrap(.~id) + scale_fill_manual(values=tbcols)+ theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.text.y = element_text(size = 14),legend.position = "right")
   rich_age_dc_plot1 
```

### Combined Age Dynamics Plot 

```{r}
#age_rich3,temporal_age_plot2,
age_grid1<-plot_grid(rich_age_dc_plot1,pc1_age_dc_plot1,nrow=2,labels="AUTO",label_size = 25)
#ggsave2('Age Dynamics Plot.pdf',age_grid1,width=28,height=12,units="cm")
#ggsave2('Age Dynamics Plot.png',age_grid1,width=25,height=25,units="cm")

```

### Stacked Barplot By Age

```{r}

# #Known Age
#   ps_age<-prune_samples(sample_data(ps_rare)$id %in% ps_all_scores3$id & !is.na(sample_data(ps_rare)$known_age),ps_rare)
# 
# #Filter to Top 5 Phyla
#   ps_age_top5phylum_filter<-subset_taxa(ps_age,Phylum %in% physeq_top5phyla)
#   
# #Plot Data
#   ps_age_top5phylum_data <- ps_age_top5phylum_filter %>%
#     microbiome::aggregate_taxa(level = "Phylum") %>%  
#     microbiome::transform(transform = "compositional") %>% plot_composition(sample.sort = "known_age") 
#   
# #Copy Across ID  
#   ps_age_top5phylum_data2<-ps_age_top5phylum_data$data
#   ps_age_top5phylum_data2$id<-sample_data(ps_age_top5phylum_filter)$id[match(ps_age_top5phylum_data2$Sample,sample_data(ps_age_top5phylum_filter)$Sample)]
#   
#    ps_age_top5phylum_data2$age<-sample_data(ps_age_top5phylum_filter)$known_age[match(ps_age_top5phylum_data2$Sample,sample_data(ps_age_top5phylum_filter)$Sample)]
#    
# #Plot
#   ps_age_top5phylum_plot <- ggplot(ps_age_top5phylum_data2, aes(x = xlabel, y = Abundance, fill = Tax)) + scale_x_discrete(breaks = ps_age_top5phylum_data2$xlabel,labels=as.character(ps_age_top5phylum_data2$age)) + geom_bar(position = "stack", stat = "identity") + facet_wrap(~id, scales = "free") + theme_bw() + scale_fill_manual(values = moma.colors("Levine2", direction = -1)) + labs(fill="Bacterial Phylum",x="Age") 
#   ps_age_top5phylum_plot
#   
#         ggsave2('Badger Barplot by Age.pdf',ps_age_top5phylum_plot,width=30,height=30,units="cm")
#         ggsave2('Badger Barplot by Age.tiff',ps_age_top5phylum_plot,width=30,height=30,units="cm")
# 
# 
# ```
# 
# ## PC1 Over Time (Date) by Infection Probability 
# Uses A different dataset filtered to known infection badgers
# ```{r}
#            
#   #Subset To Individualls with Multiuple entires 
#     woodchester_temp_table<-table(ps_all_scores2$id)
#     woodchester_scores_sub<-subset(ps_all_scores2,id %in% names(woodchester_temp_table)[woodchester_temp_table>2])
#    
#           temporal_plot1<-ggplot(woodchester_scores_sub,aes(x=sampledate_julian,y=PC1)) + geom_smooth(method="lm",se=F)+ geom_point(size=5,shape=21,aes(fill=day0_cdp),color="gray77")   
#           temporal_plot2<- temporal_plot1+ facet_wrap(.~id) + labs(x= "Julian Day",fill="M. bovis Infection Probability") + theme_bw() + theme(axis.text=element_text(size=15),strip.text.x=element_text(size=15),axis.title=element_text(size=15),legend.position = "top",axis.text.x=element_text(angle=90,hjust=1)) 
#           temporal_plot3<- temporal_plot2 + scale_fill_moma_c("Exeter")
#           temporal_plot3
#               ggsave2('Infection PC1 Temporal Plot.pdf',temporal_plot3,height=18,width=20,units="cm")
#               ggsave2('Infection PC1 Temporal Plot.tiff',temporal_plot3,height=18,width=20,units="cm")
#             
          
```


# General Linear Latent Variable Models - Continuous Tb 

```{r}

## OTUs Top 50
  badger_genus<-aggregate_top_taxa2(ps_badger_forclr, "Genus", top = 50)
  clr_scaled <-microbiome::transform(badger_genus, transform = "clr")

#Extract
  ys <- data.frame(t(otu_table(clr_scaled)))
  names(ys) <-taxa_names(clr_scaled)
  
#Predictors
  Xs<-data.frame(sample_data(clr_scaled)) %>% select(id,day0_cdp,known_age,social_group,capture_year)
  Xs$sex<-badger_rich_complete_age$sex[match(rownames(Xs),badger_rich_complete_age$Sample)]
  Xs$capture_year<-factor(Xs$capture_year)

#Define Study Design  
  sDesign<-data.frame(id = Xs$id)

## Model 
  fit_reduced_scaled <- gllvm(ys, Xs, 
                              num.lv = 3,
                              formula = ~ capture_year + day0_cdp * known_age + sex, 
                              family = "gaussian",
                              row.eff = ~(1|id),starting.val='random', studyDesign = sDesign)
  
    coefplot(fit_reduced_scaled)



#Estimates 
  df<-coef(fit_reduced_scaled)
  est_df<-data.frame(df$Intercept)
  est_df2<-data.frame(df$Xcoef) 
  est_df3<-merge(est_df, est_df2, by = 0)
  
#Order genera
  row.names(est_df3)<-est_df3$Row.names
  est_df3<-est_df3[colnames(ys),]
  names(est_df3)[1]<- "Genus"
  names(est_df3)[2]<- "Intercept"
  

### COnfidence Intervals
  confint_df<-data.frame(confint(fit_reduced_scaled))
  
#Identify Rows with Main Effects
   one_comma<-sapply(rownames(confint_df),function(x) length(gregexpr(":", x, fixed = TRUE)[[1]]))==1
   
###Strip Out Individual Datasets
  
  ##CDP
    cdp_main<-grepl("^Xcoef.day0_cdp",rownames(confint_df))
    int_data_day0cdp<-cbind(est_df3[,c("Genus","day0_cdp")],confint_df[cdp_main+one_comma==2,])
  ##Sex
    sex_main<-grepl("^Xcoef.sexM",rownames(confint_df))
    int_data_sexM<-cbind(est_df3[,c("Genus","sexM")],confint_df[sex_main,])
  ##Age
    age_main<-grepl("^Xcoef.known_age",rownames(confint_df))
    int_data_age<-cbind(est_df3[,c("Genus","known_age")],confint_df[age_main+one_comma==2,])
  #  ##Tb Sex Interaction
  #int_data_sex_day0<-cbind(est_df3[,c("Genus","sexM.day0_cdp")],confint_df[grep("Xcoef.sexM:day0_cdp",rownames(confint_df)),]) 
    #Age: Tb Interaction
  int_data_age_tb<-cbind(est_df3[,c("Genus","day0_cdp.known_age")],confint_df[grep("Xcoef.day0_cdp:known_age",rownames(confint_df)),])
  # 
 
  
#Extra Variables and Rename Columns 
  colnames(int_data_day0cdp)<-c("Genus","Estimate","l95","u95")
  colnames(int_data_sexM)<-c("Genus","Estimate","l95","u95")
  colnames(int_data_age)<-c("Genus","Estimate","l95","u95")
  colnames(int_data_age_tb)<-c("Genus","Estimate","l95","u95")
  #colnames(int_data_sex_day0)<-c("Genus","Estimate","l95","u95")

  int_data_day0cdp$trait<-"Infection Probability"
  int_data_sexM$trait<-"sex Male"
  int_data_age$trait<-"Age"
  int_data_age_tb$trait<-"Age:Infection"
  #int_data_sex_day0$trait<-"Male:Infection"
    
  tb_mod_plotdata<-rbind(int_data_day0cdp,int_data_sexM,int_data_age,int_data_age_tb)

#Order   
  tb_mod_plotdata$trait<-factor(tb_mod_plotdata$trait,levels=c("sex Male","Age","Infection Probability","Age:Infection"))
  tb_mod_plotdata2<- tb_mod_plotdata %>% group_by(trait)  %>% arrange(Estimate,.by_group=T)
  tb_mod_plotdata2$Genus<-factor(tb_mod_plotdata2$Genus,levels=unique(tb_mod_plotdata2$Genus))
#Significance  
  tb_mod_plotdata2$Sig<- !data.table::between(0, tb_mod_plotdata2$l95, tb_mod_plotdata2$u95)

   sig_col<-moma.colors("Levine2")[6]

##Subset
   tb_mod_plotdata_sigtab<-with(tb_mod_plotdata2,table(Genus,Sig))
   tb_mod_plotdata_sigsubset<-subset(tb_mod_plotdata2,Genus %in% rownames(tb_mod_plotdata_sigtab)[tb_mod_plotdata_sigtab[,2]>0])
   
### Plot
  tb_plot1<-ggplot(tb_mod_plotdata_sigsubset,aes(x=Estimate,y=Genus)) + geom_errorbarh(aes(xmin=l95,xmax=u95,color=Sig),linewidth = 1.2,alpha=0.7) + geom_point(size=7,shape=21,color="gray40",aes(fill=Sig),alpha=0.7)
  tb_plot2<- tb_plot1 + theme_bw(base_size = 20) + geom_vline(xintercept=0,linetype="dashed") + scale_color_manual(values=c("gray77",sig_col)) + scale_fill_manual(values=c("white",sig_col)) + guides(fill="none",color="none") + facet_wrap(.~trait,scales = "free_x", nrow = 1) + 
    theme(
    axis.text = element_text(size = 22),
    axis.title = element_text(size = 26),
    strip.text = element_text(size = 25))
  tb_plot2
  
ggsave('Age Infection GLLVM plot.png', tb_plot2, width = 45, height = 20,units="cm")
  
#### Correlations
  cr1<-data.frame(getResidualCor(fit_reduced_scaled))#
  
  names(cr1)<-names(ys)
  names(cr1)<-abbreviate(names(cr1), minlength = 15)
  rownames(cr1)<-abbreviate(rownames(cr1), minlength = 15 )
  
  library(rstatix)
  #devtools::install_github("kassambara/ggcorrplot")
  library(ggcorrplot)
  #install.packages("ggpubr")
  library(ggpubr)
  
  cr2<-cor_pmat(cr1)
  
  corplot<-ggcorrplot(cr1, 
                      hc.order = TRUE,
                      outline.col = "white",
                      type = "full",
                      ggtheme = ggplot2::theme_minimal(base_size = 10),
                      tl.cex = 12,
                      p.mat = cr2,
                      sig.level = 0.05,
                      lab_size = 30,
                      #show.diag = F,
                      insig = "blank",
                      # colors = c("#6D9EC1", "white", "#E46726"))
                      colors = c("blue", "white", "red"))+
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 20), axis.text.y = element_text(size = 20),legend.text = element_text(size=20),legend.title = element_text(size=25),legend.key.size = unit(1, 'cm'))+
    theme(plot.margin=unit(c(0.2,0.2,0.2,2),"cm"))
  corplot
  
ggsave('Correlation plot.png', corplot, width = 40, height = 40,units="cm")
  
## Save Grid
  #tb_mod_combined1<-ggarrange(tb_plot2,corplot,nrow=2,heights=c(0.8,1),widths=c(1,1),align="h",labels="AUTO",font.label=list(size=50))
  #tb_mod_combined1
   # cowplot::ggsave2('Tb Combined Model Outputs.png',tb_mod_combined1,width=20,height=25)
  #  cowplot::ggsave2('Tb Combined Model Outputs.pdf',tb_mod_combined1,width=20,height=25)


```

# General Linear Latent Variable Models - Categorical Tb 

```{r}

## OTUs Top 50
  ps_badger_gllvmdata<-subset_samples(ps_badger_forclr,!is.na(sample_data(ps_badger_forclr)$btb_status_time))
  badger_gllvmdata_genus<-aggregate_top_taxa2(ps_badger_gllvmdata, "Genus", top = 50)
  clr_scaled_tbcat <-microbiome::transform(badger_gllvmdata_genus, transform = "clr")

#Extract
  y_cattb <- data.frame(t(otu_table(clr_scaled_tbcat)))
  names(y_cattb) <-taxa_names(clr_scaled_tbcat)
  
#Predictors
  X_cattb<-data.frame(sample_data(badger_gllvmdata_genus)) %>% select(id,btb_status_time,known_age,social_group,capture_year)
  X_cattb$sex<-badger_rich_complete_age$sex[match(rownames(X_cattb),badger_rich_complete_age$Sample)]
  X_cattb$capture_year<-factor(X_cattb$capture_year)
  X_cattb$btb_status_time<-factor(X_cattb$btb_status_time,levels=c("Unexposed","Exposed","Excreter"))
  
#Define Study Design  
  sDesign_cattb<-data.frame(id = X_cattb$id)

## Model 
  fit_cattb1 <- gllvm(y_cattb, X_cattb, 
                              num.lv = 2,
                              formula = ~ capture_year + btb_status_time * known_age + sex, 
                              family = "gaussian",
                              row.eff = ~(1|id),starting.val='random', studyDesign = sDesign_cattb)
  
  #coefplot(fit_cattb1)
```

## Plots of Genera by Predictor

### Age
```{r}

badger_gllvmdata_genus
otu_table(badger_gllvmdata_genus)

#Fusobacterium Age 
  fuso_df<-data.frame(count=as.numeric(otu_table(clr_scaled)[which(rownames(otu_table(clr_scaled))=="Fusobacterium"),]),age=sample_data(badger_genus)$known_age)
  
  fusoplot1<-ggplot(fuso_df,aes(x=age,y=count)) + geom_smooth(method="lm") + geom_point(fill= moma.colors("Andri",3)[3],alpha=0.7,shape=21,size=5,position = position_jitter(width=0.1)) + theme_bw(base_size=15) + labs(x="Badger Age",y="Rel. Abundance \n (CLR-Transformed") + scale_x_continuous(breaks = seq(0,10,1)) + geom_text(x=8,y=-2.5,label="Fusobacterium",cex=7)
fusoplot1

#Morganella Age 
  morganella_df<-data.frame(count=as.numeric(otu_table(clr_scaled)[which(rownames(otu_table(clr_scaled))=="Morganella"),]),age=sample_data(badger_genus)$known_age)
  
  morganplot1<-ggplot(morganella_df,aes(x=age,y=count)) + geom_smooth(method="lm") + geom_point(fill= moma.colors("Andri",1),shape=21,size=5,position = position_jitter(width=0.1),alpha=0.7,colour="white") + theme_bw() + labs(x="Badger Age",y="Morganella Rel. Abundance \n (CLR-Transformed") + scale_x_continuous(breaks = seq(0,10,1)) 
  morganplot1
  
#Ligilactobacillus Age 
  ligil_df<-data.frame(count=as.numeric(otu_table(clr_scaled)[which(rownames(otu_table(clr_scaled))=="Ligilactobacillus"),]),age=sample_data(badger_genus)$known_age)
  
  ligilplot1<-ggplot(ligil_df,aes(x=age,y=count)) + geom_smooth(method="lm") + geom_point(fill= moma.colors("Andri",1),alpha=0.7,shape=21,size=5,position = position_jitter(width=0.1)) + theme_bw(base_size=15) + labs(x="Badger Age",y="Rel. Abundance \n (CLR-Transformed") + scale_x_continuous(breaks = seq(0,10,1)) + geom_text(x=8,y=7,label="Ligilactobacillus",cex=7)
  ligilplot1
  
#Lactobacillus Infection 
  lacto_df<-data.frame(count=as.numeric(otu_table(clr_scaled)[which(rownames(otu_table(clr_scaled))=="Lactobacillus"),]),age=sample_data(badger_genus)$known_age,infection=sample_data(badger_genus)$day0_cdp)
  
  lactoplot1<-ggplot(lacto_df,aes(x=infection,y=count)) + geom_smooth(method="lm") + geom_point(fill= moma.colors("Flash",1),alpha=0.7,shape=21,size=5,position = position_jitter(width=0.1)) + theme_bw(base_size=15) + labs(x="M. bovis Infection Probability",y="Rel. Abundance \n (CLR-Transformed") + scale_x_continuous(breaks = seq(0,1,0.1)) + geom_text(x=0.8,y=3,label="Lactobacillus",cex=7)
  lactoplot1
  
```

### Age: Infection
  
```{r}  
  
#Bacteroides 
  bacteroides_df<-data.frame(count=as.numeric(otu_table(clr_scaled)[which(rownames(otu_table(clr_scaled))=="Bacteroides"),]),age=(sample_data(badger_genus)$known_age),infection=sample_data(badger_genus)$day0_cdp)

#Model to Pull Out Interaction
  bacmod1<-glm(count ~ age*infection,data=bacteroides_df)
  bac_newdf<-data.frame(infection=rep(c(0,1),each=10),age=rep(seq(10),2))
  bac_newdf$count<-predict(bacmod1,newdata = bac_newdf)
  
  bacteroidesplot1<-ggplot(data=bacteroides_df,aes(x=age,y=count)) + geom_point(aes(fill=infection),shape=21,size=5,position = position_jitter(width=0.1)) + theme_bw(base_size=15) +  scale_fill_moma_c("Exter") + labs(fill="M. bovis \nInfection Probability",x="Badger Age",y="Relative Abundundance \n (CLR-Transformed)")
bacteroidesplot2<-bacteroidesplot1 + geom_line(data=bac_newdf,aes(x=age,y=count,group=factor(infection),colour = factor(infection)),cex=2) + scale_color_manual(values=moma.colors("Exter",5)[c(1,5)]) + guides(color=FALSE) + theme(legend.position = "top") + scale_x_continuous(breaks = seq(0,10,1)) + geom_text(x=8,y=-1,label="Bacteroides",cex=7)
bacteroidesplot2  

  #Clostridium 
  clost_df<-data.frame(count=as.numeric(otu_table(clr_scaled)[which(rownames(otu_table(clr_scaled))=="Clostridium sensu stricto 13"),]),age=(sample_data(badger_genus)$known_age),infection=sample_data(badger_genus)$day0_cdp)
  
#Model to Pull Out Interaction
  clostmod1<-glm(count ~ age*infection,data=clost_df)
  clost_newdf<-data.frame(infection=rep(c(0,1),each=10),age=rep(seq(10),2))
  clost_newdf$count<-predict(clostmod1,newdata = clost_newdf)

  clostplot1<-ggplot(clost_df,aes(x=age,y=count)) + geom_point(shape=21,size=5,aes(fill=infection),position = position_jitter(width=0.1)) + theme_bw(base_size=15) +  scale_fill_moma_c("Exter") + labs(fill="M. bovis \nInfection Probability",x="Badger Age",y="Relative Abundance \n (CLR-Transformed)") + geom_text(x=6,y=6,label="Clostridium \nSensu Stricto 13",cex=7)
 clostplot2<-clostplot1 + geom_line(data=clost_newdf,aes(x=age,y=count,group=factor(infection),colour = factor(infection)),cex=2) + scale_color_manual(values=moma.colors("Exter",5)[c(1,5)]) + guides(color=FALSE) + theme(legend.position = "top") + scale_x_continuous(breaks = seq(0,10,1)) 
clostplot2
 
 
 
 
 #Staph 
  staph_df<-data.frame(count=as.numeric(otu_table(clr_scaled)[which(rownames(otu_table(clr_scaled))=="Staphylococcus"),]),age=factor(sample_data(badger_genus)$known_age),infection=sample_data(badger_genus)$day0_cdp)
  
  staphplot1<-ggplot(staph_df,aes(x=age,y=count)) + geom_smooth(method="lm") + geom_point(shape=21,size=5,aes(fill=infection),position = position_jitter(width=0.1)) + theme_bw(base_size=15) +  scale_fill_moma_c("Exter") + labs(fill="M. bovis \nInfection Probability",x="Badger Age",y="Staphylococcus Rel. Abund. \n (CLR-Transformed)")
 staphplot1
     
#Salmonella 
  salmonella_df<-data.frame(count=as.numeric(otu_table(clr_scaled_tbcat)[which(rownames(otu_table(clr_scaled_tbcat))=="Salmonella"),]),age=sample_data(badger_gllvmdata_genus)$known_age)
  
  ggplot(salmonella_df,aes(x=age,y=count)) + geom_smooth(method="lm") + geom_point(shape=21,size=5,fill="lightblue",position = position_jitter(width=0.1)) + theme_bw()   
```

## Combined Plot

```{r}

gllvmplot1<-plot_grid(fusoplot1,ligilplot1,lactoplot1,bacteroidesplot2,clostplot2,nrow=2,ncol=3,labels="AUTO",label_size=20)
gllvmplot1
ggsave2('Badger Genus Trends.png',gllvmplot1,width=40,height = 25,units="cm")
```
